---
title: "Geometric distribution"
output:
  beamer_presentation
date: October 1, 2016
author: Timothy Daley
---



# Geometric distribution {.bigger}

# Geometric distribution


- Suppose you flip a coin and count the number of times you need to flip until you get a heads.  What is the probability that you will have to get $x$ tails to get a result of heads? (We don't count the last flip since it will be heads)


- $\Pr(x \text{ tails}) = \Pr((x) \text{ tails}) \cdot \Pr(\text{1 heads})$ $$= (1 - p)^{x} p$$

- Note that we differ from the book here, see [https://en.wikipedia.org/wiki/Geometric_distribution] for an explanation of both.

# Geometric distribution

- $X \sim \text{Geometric}(p)$, $p =$ probability of success


- pmf: $f(x) = \Pr(X = x) = p (1 - p)^{x}$, $x \geq 0$


- cdf: $F(x) = \Pr(X \leq x) = \sum_{i = 0}^{x} f(x) =$ $\sum_{i = 0}^{x} (1 - p)^{x} p = 1 - (1 - p)^{x + 1}$


# Properties of the Geometric distribution


- Expectation: $\mathrm{E}(X) = \frac{1 - p}{p} = \frac{1}{p} - 1$


- Variance: $\mathrm{Var}(X) = \frac{1 - p}{p^{2}}$


# Geometric distribution in R

R uses the parametrization that $X$ is the number of failures before the first success, which agrees with our parametrization.  Remember that if you're counting the total number you will need to add 1.

- pmf: $f(x) =$ dgeom(x, prob = $p$)

- cdf: $F(x) = \Pr(X \leq x) =$ pgeom(x, prob = $p$)

- random number generation: $x =$ rgeom(n, prob = $p$)

# Examples: designing a survey

Suppose you want to design a survey and need at least one person representative of each group.  In California, the proportion of black non-Hispanic residents is $5.8 \%$ (http://www.pewresearch.org/fact-tank/2014/01/24/in-2014-latinos-will-surpass-whites-as-largest-racialethnic-group-in-california/).  What is the probability that you will need to interview at least 20 residents to include at least 1 black non-Hispanic resident?

# Examples: designing a survey

- $X =$ \# residents interviewed until the first black non-Hispanic resident.

- $X - 1 \sim \text{Geometric}(0.058)$

- $\Pr(X - 1 \geq 20) =  \Pr(X \geq 19) = 1 - \Pr(X \leq 18)$ 
    ```{r}
    1 - pgeom(18, prob = 0.058)
    ```
    
# Examples: designing a survey

- How many people should you expect to interview such that the probability of selecting at least 1 black non-Hispanic resident is $95 \%$?

- Find first $x$ such that $F(x) \geq 0.95$.  Know that $F(19) = 0.68$
    ```{r}
    i = 19
    F = pgeom(i, prob = 0.058)
    while(F < 0.95){
      i = i + 1
      F = pgeom(i, prob = 0.058)
    }
    i + 1 # add 1 to total
    ```


    
# Examples: bit parsing

- Consider a digital transmission of bits.  The bits are transmitted until the first error and each bit can independently be an error with a $1 \%$ probability.  What is the probability that at least 100 bits will be transmitted before failure?

- Total number of bits = $X + 1$ = successful bits + failed bit.

- $X \sim \text{Geometric}(p = 0.01)$

- $\Pr(X \geq 100) = 1 - \Pr(X \leq 99)$
    ```{r}
    1 - pgeom(99, prob = 0.01)
    ```
    
# Previous example: manufacturing

In a previous example, a tent manufacturer's track record indicates that it averages 1 fault per 50 square feet of canvas.  What is the average number of faultless 75 ft$^{2}$ tents that it will make till it produces one with a fault?

- $\Pr(\text{tent with at least failures}) = 1 - e^{- 1.5} = p$

- Number of tents without failures = $X \sim \text{Geometric}(p = 1 - e^{- 1.5})$  

- Expected \# tents without failures = $\mathrm{E}(X) = 1/p - 1 = 1/(1 - e^{-1.5}) \approx 0.29$.

# Memoryless property of the geometric distribution

What if we know that we have gone 5 tents without a failure.  How does this change the distribution?

- $\Pr(X = x | X \geq 5) = \frac{\Pr(X = x \cap X \geq 5)}{\Pr(X \geq 5)}$ $$= \frac{(1 - p)^{x} p}{(1 - p)^{5}} = (1 - p)^{x - 5}p$$ $$= \Pr(X = x - 5)$$

- When we condition on the number of failures already observed, it's like we start over now.  Hence, the memoryless property of the geometric distribution.

