---
title: "Homework 1 Solutions"
output: pdf_document
---

1. Chapter 2, excercise 1
    Q. This problem concerns the ALOHA network model of Section 2.1. Feel free to use (but cite) computations already in the example.
        (a) $\Pr(X_{1} = 2 and X_{2} = 1)$, for the same values of $p$ and $q$ in the examples. 
        (b) Find $\Pr(X_{2} = 0)$.
        (c) Find $\Pr(X_{1} = 1|X_{2} = 1)$.
    A.  $p = 0.4$ and $q = 0.8$
      a.  From the example, $X_{1} = 2$ only when both nodes are active and try to send or they don't try to send.  If both nodes try to send, then there is collision and are both active again at time 2, i.e. $X_{2} = 2$.  If both nodes don't try to send, then $X_{2} = 1$ if node sends and the the other refrains.  This occurs with probability $(1 - p)^{2} \cdot 2 p (1 - p)$.
      ```{r cache=TRUE}
      p = 0.4;
      q = 0.8;
      ((1 - p)^2)*2*p*(1 - p)
      ```
      b. $X_{2} = 0$ only if one node transmits in time 1 and does not generate a new message at time 2, while the other node refrains at time 1 and sends a message at time 2.  Since there are two ways for this to happen, the probability is equal to $2 \cdot p \cdot (1 - q) \cdot (1 - p) \cdot p$.
      ```{r}
      2*p*(1 - q)*(1 - p)*p
      ```
      c. Note that $\Pr(X_{1} = 1 | X_{2} = 1) = \Pr(X_{1} = 1 \cap X_{2} = 1) \cdot \Pr(X_{2} = 1)$.  $(X_{1} = 1 \cap X_{2} = 1)$ only occurs if one node transmits at time 1 and then remains inactive at time 2 while the other does not transmit at time 1 and 2.  The probability of this is
      ```{r}
      2*(p*(1 - q)*(1 - p)^2)
      ```
      \newline 
      Let us now think about how $X_{2} = 0$ can occur in other ways.  We can have both nodes active at time 1, and then only one is active at time $2$.  This occurs with probability
      ```{r}
      (p^2 + (1 - p)^2)*(2*p*(1 - p))
      ```
      \newline
      Therefore $\Pr(X_{1} = 1 | X_{2} = 1) = 0.0576/(0.0576 + 0.2496) =$
      
2. Chapter 2, excercise 2
    a.  There are two cases: we draw a blue marble from urn one and we draw a yellow marble from urn one.  If we draw a blue marble (with probability $3/6$), then the probability of drawing a blue marble from urn 2 is $(5 + 1)/13$. If we draw a yellow marble from urn 1 (with probability $1/2$), then the probability of drawing a blue marble from urn 2 is $5/(13)$.  Therefore the total probability is 
    ```{r}
    (3/6)*(6/13) + (3/6)*(5/13)
    ```
    b. The probability the first marble is blue given the second marble is blue is the first of the above probabilities divided by the above.
    ```{r}
    (3/6)*(6/13)/((3/6)*(6/13) + (3/6)*(5/13))
    ```


3. Chapter 2, excercise 4
    The number of ways that two out of five programs can fail is $\binom{5}{2} = 10$ and each occurs with probability $0.2^2*(1 - 0.2)^3$, giving a total probability of 
```{r}
choose(5, 2)*0.2^2*(1 - 0.2)^3
```

4. Chapter 2, excercise 6
    a.  Note first that $\Pr(\text{at least one prize}) = 1 - \Pr(\text{no prizes})$.  The probability of winning no prizes is equal to the probability that none of the five tickets drawn match your three, leaving 57 numbers to choose from.  Therefore the number of ways to do this is $\binom{57}{5}$ out of $\binom{60}{5}$ total ways.
    ```{r}
    1 - choose(57, 5)/choose(60, 5)
    ```
    b. The probability that you win exactly one prize is the probability that exactly one of the five drawn number matches one of your three.  The number of ways this can occur is $\binom{3}{1} \cdot \binom{57}{4}$.
    ```{r}
    choose(3, 1)*choose(57, 4)/choose(60, 5)
    ```

5. Chapter 2, excercise 12 (a and b only)
    a. The probability of winning is $\Pr(N = 3 | N \geq 3)$.  Note that $3 \leq N \leq 5)$.  You can get $(N = 3)$ 3 ways:  3 on first roll, 1 on first roll and 2 on second, 2 on first roll and 1 on second, and finally 1 on first 3 rolls.  The probability of these are $1/3$, $1/9$, $1/9$, and $1/27$, respectively.  To get 4, one can roll a 2 first roll and a 2 the second roll, a 1 the first roll and 3 the second roll, or a 1 the first two rolls and a 2 the third roll.  The probabilities are $1/9$, $1/9$, and $1/27$.  To get a 5 you can roll a 2 the first roll and three the second or a 1 the first two rolls and 3 the second.  The probability of these are $1/9$ and $1/27$.  Therefore $\Pr(N = 3 | N \geq 3) =$
    ```{r}
    (1/3 + 1/9 + 1/9 + 1/27)/(1/3 + 1/9 + 1/9 + 1/27 + 1/9 + 1/9 + 1/27 + 1/9 + 1/27)
    ```
    b. Using the above calculated outcomes, if we won then the first roll was a 1 if we get a 1 on the first roll and 2 on the second or a 1 on all three rolls.  Therefore $\Pr(\text{first roll is 1} | \text{win}) =$
    ```{r}
    (1/9 + 1/27)/(1/3 + 1/9 + 1/9 + 1/27)
    ```
    
6. Chapter 2, excercise 14
    a.  The probability that the number of turns is one is the probability that either Jack or Jill shows a 1 and the other does not.  This equals
    ```{r}
    (1/4)*(2/3) + (3/4)*(1/3)
    ```
    In order for the number of turns to equal 2, first it must not equal 1.  Then either must roll a 1 but not both.  This occurs with probability
    ```{r}
    (1 - (1/4)*(2/3) + (3/4)*(1/3))*((1/4)*(2/3) + (3/4)*(1/3))
    ```
    b. If $N = 2$, then in the first turn  one of two things happened: neither rolled a 1 (with probability $(3/4)*(2/3)$), or both rolled a 1 (with probability $(1/4)*(1/3)$).  Therefore $\Pr(\text{first turn was a tie} | N = 2) =$
    ```{r}
    (1/4)*(1/3)/((1/4)*(1/3) + (3/4)*(2/3))
    ```