---
title: "Quiz 8 Solutions"
author: "Name: "
output: pdf_document
---

Remember to state the logic behind your answers.  If you use any variables, please state what the variables represent.

1.  Which of the following plots show signs of heteroscedasticity?

```{r echo=FALSE, fig.width=8, fig.height=3, fig.align='center'}
par(mfrow = c(1, 4))
n_obs = 80
x = runif(n_obs, 0, 10)
y = 2*x + 3 + rnorm(n_obs, 0, 5)
plot(x, y, type = "p", pch = 16, main = "a")
abline(lm(y ~ x))
x = runif(n_obs, 0, 10)
y = 2*x + 3 + rnorm(n_obs, 0, x)
plot(x, y, type = "p", pch = 16, main = "b")
abline(lm(y ~ x))
x = runif(n_obs, 0, 10)
y = exp(-x + rnorm(n_obs, 0, 1))
plot(x, y, type = "p", pch = 16, main = "c")
abline(lm(y ~ x))
x = runif(n_obs, 0, 10)
y = rnorm(n_obs, 0, 8)
plot(x, y, type = "p", pch = 16, main = "d")
abline(lm(y ~ x))
```

b and c.

2.  From the below regression summary answer the following questions.
```{r echo = FALSE}
library(ISLR)
data(Auto)
library(pander)
panderOptions("digits", 3)
```
```{r}
credit = read.csv(file = "http://www-bcf.usc.edu/~gareth/ISL/Credit.csv")
income_vs_education.lm = lm(Income ~ Education, data = credit)
summary(income_vs_education.lm)
```
\newpage
a.  What is the $R^{2}$ of this regression and what is correlation between the fitted income and the observed income?  

$R^{2} = 0.0007668$.  The correlation between the fitted values and the observed is equal to the square root of the $R^{2}$, which equals $0.0277$

b.  Is the coefficient for Education significantly different from 0 at the $95 \%$ confidence level?

No.  $p = 0.581$.

c.  Assuming the linear model is correct, an increase in 1 year of Education corresponds to what expected increase of Income?

$-0.3123$.

d.  Below are diagnostic plots for the above regression.  Does the fit of this regression look good and why or why not?

No.  It looks terrible based on the plot and the QQ plot.  It looks like theres a ton of variation in income that is unexplained by Education.

```{r fig.height=2.8, fig.width=6, fig.align='center'}
plot(x = credit$Education, y = credit$Income, type = "p", pch = 16)
abline(income_vs_education.lm, lwd = 2)
```
```{r fig.height=2.5, fig.width=8, fig.align='center'}
par(mfrow = c(1, 4))
plot(income_vs_education.lm)
```
