---
title: "Quiz 8"
author: "Name: "
output: pdf_document
---

Remember to state the logic behind your answers.  If you use any variables, please state what the variables represent.

1.  Which of the following plots show signs of heteroscedasticity?

```{r echo=FALSE, fig.width=8, fig.height=3, fig.align='center'}
par(mfrow = c(1, 4))
n_obs = 80
x = runif(n_obs, 0, 10)
y = 2*x + 3 + rnorm(n_obs, 0, 5)
plot(x, y, type = "p", pch = 16, main = "a")
abline(lm(y ~ x))
x = runif(n_obs, 0, 10)
y = 2*x + 3 + rnorm(n_obs, 0, x)
plot(x, y, type = "p", pch = 16, main = "b")
abline(lm(y ~ x))
x = runif(n_obs, 0, 10)
y = exp(-x + rnorm(n_obs, 0, 1))
plot(x, y, type = "p", pch = 16, main = "c")
abline(lm(y ~ x))
x = runif(n_obs, 0, 10)
y = rnorm(n_obs, 0, 8)
plot(x, y, type = "p", pch = 16, main = "d")
abline(lm(y ~ x))
```

2.  From the below regression summary answer the following questions.
```{r echo = FALSE}
library(ISLR)
data(Auto)
library(pander)
panderOptions("digits", 3)
```
```{r}
credit = read.csv(file = "http://www-bcf.usc.edu/~gareth/ISL/Credit.csv")
income_vs_education.lm = lm(Income ~ Education, data = credit)
summary(income_vs_education.lm)
```
\newpage
a.  What is the $R^{2}$ of this regression and what is correlation between the fitted income and the observed income?  
\
\
\
b.  Is the coefficient for Education significantly different from 0 at the $95 \%$ confidence level?
\
\
\
c.  Assuming the linear model is correct, an increase in 1 year of Education corresponds to what expected increase of Income?
\
\
\
\
d.  Below are diagnostic plots for the above regression.  Does the fit of this regression look good and why or why not?
\
\

```{r fig.height=2.8, fig.width=6, fig.align='center'}
plot(x = credit$Education, y = credit$Income, type = "p", pch = 16)
abline(income_vs_education.lm, lwd = 2)
```
```{r fig.height=2.5, fig.width=8, fig.align='center'}
par(mfrow = c(1, 4))
plot(income_vs_education.lm)
```
