---
title: "Quiz 5"
author: "Name: "
output: pdf_document
---

1.  Suppose that $x_{1}, \ldots, x_{n}$ are independent and identically distributed observations from a normal distribution with mean $10$ and variance $4$.  What is the distribution of $\bar{x}$ for $n = 10$?
   \
    \
    \
    \
    \
    \
    \
    \
    \
    \
2.  Suppose that we measure want to compare a new flu treatment to the traditional treatment.  The average recovery time for the traditional treatment is 4 days.  To test the new treatment you give it to 50 people and obtain an average recovery time of 2.8 days with a standard deviation of 1.9 days.
    a. What is the standard error of the mean?
    \
    \
    \
    \
    \
    \
    \
    b. Construct a $95 \%$ confidence interval for the mean recovery time of the new treatment.
    \
    \
    \
    \
    \
    \
    \
    \
    \
    \
    b. Approximately what sample size would you need to obtain a margin of error of $0.5$?

\pagebreak

Standard normal table
```{r echo=FALSE}
library(knitr)
x = data.frame(alpha = c(0.1, 0.05, 0.01), z = c(qnorm(1 - 0.1/2), qnorm(1 - 0.05/2), qnorm(1 - 0.01/2)))
x
#kable(x, format = "markdown")
```

$t$ table
```{r echo=FALSE}
library(knitr)
x = data.frame(alpha = rep(c(0.1, 0.05, 0.01), times = 9), df = rep(2:10, each = 3))
x = data.frame(alpha = x$alpha, df = x$df, t = qt(1 - x$alpha/2, df = x$df) )
x
#kable(x, format = "markdown", colnames = colnames(x))
```
    